---
title: "Chapter 8: Non-parametric test and assumptions"
description: "Learn about non-parametric test and assumptions"
date: "05/21/2022"
format: html
categories: [statistics, basic statistics]
draft: true
---

```{r}
#| include: false
paged_print <- function(x, options) {
  knitr::asis_output(
    rmarkdown:::paged_table_html(x, options = attr(
      x,
      "options"
    )),
    meta = list(dependencies = rmarkdown::html_dependency_pagedtable())
  )
}

registerS3method(
  "knit_print", "data.frame", paged_print,
  envir = asNamespace("knitr")
)
```

::: {layout-ncol="1"}
![Which is more robust between paramteric and non-paramteric? Image source: [Twitter](https://twitter.com/v_matzek/status/1376959858042306562)](images/para.jpg){.preview-image} Which is more robust between paramteric and non-paramteric?

Image source: [Twitter](https://twitter.com/v_matzek/status/1376959858042306562)
:::

## Non-parametric test and assumptions

Non-parametric tests are used when the assumptions for parametric tests fail;

1.  The data does not normal distribution
2.  The groups should have un-equal variances (heteroskedasticity)
3.  The two groups are not independent of each other

## Mann--Whitney U test

### one sample mann-whitney u test

The `Mann–Whitney U test` also known as the `Mann-Whitney-Wilcoxon test` is the equivalent of one sample and two-sample t-test for non-normal data. Like last time, in this tutorial also, we will be using the `penguins` dataset from the `{palmerpenguins}` package in R. For the sake of the tutorial, we will assume that the assumptions for parametric tests are violated here.

We will we checking if the median bill length of Gentoo penguins is significantly different from 50mm.

```{r}
#install.packages("palmerpenguins")
#install.packages("tidyverse")
library(palmerpenguins)
library(tidyverse)

pen_gen <- penguins %>% filter(species=="Gentoo")
# one sample mann-whitney test
wilcox.test(pen_gen$bill_length_mm, mu = 50)
```

### two sample mann-whitney u test

We will we checking if bill lengths of Gentoo penguins differ significantly from the bill length of Adelie. This will be a two sample test.

```{r}
#install.packages("palmerpenguins")
#install.packages("tidyverse")
library(palmerpenguins)
library(tidyverse)

pen_gen <- penguins %>% filter(species=="Gentoo")
pen_ad <- penguins %>% filter(species=="Adelie")
# two sample mann-whitney test
wilcox.test(pen_gen$bill_length_mm, pen_ad$bill_length_mm)
```

### wilcoxon signed rank test

The Wilcoxon Signed-Rank test is the equivalent of paired t-test.

We will be checking if there is a significant difference between the bill length and flipper length in Chinstrap penguins.

```{r}
#install.packages("palmerpenguins")
#install.packages("tidyverse")
library(palmerpenguins)
library(tidyverse)

pen_chin <- penguins %>% filter(species=="Chinstrap")
# wilcoxon singed rank test
wilcox.test(pen_chin$bill_length_mm, pen_chin$flipper_length_mm, paired = T)
```

## References

1.  Neuhäuser M. (2011) Wilcoxon--Mann--Whitney Test. In: Lovric M. (eds) International Encyclopedia of Statistical Science. Springer, Berlin, Heidelberg. https://doi.org/10.1007/978-3-642-04898-2_615 [Link](https://link.springer.com/referenceworkentry/10.1007/978-3-642-04898-2_615)

#### Last updated on {.unnumbered .unlisted}

```{r}
#| echo: false
Sys.time()
```

<a hidden href="https://info.flagcounter.com/ynrK"><img src="https://s11.flagcounter.com/count2/ynrK/bg_000000/txt_FFFFFF/border_F0F0F0/columns_5/maxflags_25/viewers_0/labels_1/pageviews_1/flags_0/percent_0/" alt="Flag Counter" border="0"/></a>
